\documentclass[itdr/core]{subfiles}
\pdfoutput=1
\begin{document}

\toc[2] % Table of Contents

\vfill

\noindent
\begin{minipage}{\textwidth}
\begin{center}\fnsize
	Writing and illustrations by \author{}.

	``\title'' is licensed under\\the \license.
\end{center}
\end{minipage}

\clearpage

\chapterx{Rules}
\label{ch:Introduction}
\section{Die roll convention.}
The game uses common six-sided dice, eight-sided dice, ten-sided dice and twelve sided dice. Each die you need to roll
is noted as ``d''. Unless otherwise noted otherwise, ``2d'', for example means \emph{roll two dice and add the result
together}. 

\section{The Basic Game Mechanic}

The basic game mechanic is the check. You throw 2d6, add the relevant skill and relevant modifier. A \emph{natural}
result of $\mathbf{2}$ is always a failure, regardless of skills and modifiers. A \emph{natural} result of $\mathbf{12}$ is always a success, even in adverse situations. 

\begin{longtable}{p{0.1\columnwidth}p{0.8\columnwidth}}
\hline
& \textbf{Skill Check Difficulties} \\
\hline
6 & A relatively simple task that is still more than the PC would usually be expected to manage in their regular background. Anything easier than this isn't worth a skill check. \\
8 & A significant challenge to a competent professional that they'd still suceed at more often than not. \\
10 & Something too difficult to be expected of anyone but a skilled expert, and even they might fail. \\
12 & Only a true master could expect to carry this off with any degree of reliability. \\
14+ & Only a true master could has any chance of achieving this at all, and even they will probably fail. \\
\hline
\caption{Skill Check Difficulty Table}
\label{table:skill-checks}
\end{longtable}

\section{Skills} 
Your hero's learned capabilities are skills, and every character has their share of them. Skills are measured in a rating from level-0 for competent practitioners and level-5 for one of the best in the world. Novice characters start with level-0 or level-1 in their skills.

\begin{itemize}
\item \textbf{Animals}: the care and training of animals. Also covers general agriculture.

\item \textbf{Admin}:  keep an organization running smoothly, scribe things well, plan out logistics, identify incompetent or treacherous workers, analyze records or archives, or otherwise do things that an executive or middle-manager would need to do.

\item \textbf{Connect}: find or know people who are useful to your purposes, make friendships or social acquaintances, know who to talk to get favors or services, and call on the help or resources of organizations you belong to. Connect covers your PC’s ability to find the people you need, though convincing them to help may require more than this.

\item \textbf{Craft}: the ability to maintain, repair, and build mechanical devices of all sorts, from crossbows to cartwheels. Also covers lockpicking.

\item \textbf{Exert}: the ability to exert oneself physically. You may add the Athletics skill to appropriate physical characterisc rolls. This skill is never rolled unskilled.

\item \textbf{Heal}: treat wounds, cure diseases, neutralize poisons, diagnose psychological health issues, and otherwise tend to the wounds of body and mind. The Heal skill cannot cure lost hit points directly, but it’s a vital skill in stabilizing Mortally Wounded allies or ensuring clean recovery from grievous injuries.

\item \textbf{Deception}: convincingly avoiding the truth and misleading other people.

\item \textbf{Investigation}: this skill combines keen observation, forensics, research, and detailed analysis.

\item \textbf{Know}: know matters of history, geography, natural science, zoology, and other academic fields appropriate to a sage or scholar. While some sages might specialize in particular fields, most learned men and women in this age have a broad range of understanding, and will rarely be unable to even attempt to answer a question relevant to this skill.

\item \textbf{Lead}: inspire others to follow your lead and believe in our plans and goals. Manage subordinates and keep them focused, loyal, and motivated in the face of danger or failure. A successful leader will keep their subordinate’s faith and confidence even when reason might make the leader’s plan appear questionable at best.

\item \textbf{Magic}: cast or analyze magic and know things about famous mages or notable magical events. Classes that can’t cast spells obtain only intellectual and scholarly benefits from this skill.

\item \textbf{Notice}: notice small details, impending ambushes, hidden features, or concealed objects. Detect subtle smells, sounds, or other sensory input. Notice cannot be used simply to detect a lie, but keen attention can often discern a subject’s emotional state.

\item \textbf{Liaison}: the art and practice of negotiation and diplomacy in a myriad of social situations. This skill covers diplomatic meetings, haggling with merchants, or diffusing a situation before blades start flashing.

\item \textbf{Perform}: sing, act, dance, orate, or otherwise perform impressively for an audience. Compose music, plays, writings, or other works of performance art. Most performers will have a particular field they excel at, though polymaths might exist if the PC’s background is appropriate for such versatility.

\item \textbf{Pray}: perform the clerical rites of your religion, and be familiar with the gods, demons, and taboos of major and minor faiths, and identify iconography and persons of religious importance. Pray also helps you know the state of local faiths and the important persons in their hierarchies.

\item \textbf{Punch}: fight unarmed or with natural body weaponry. Punch, kick, grapple, or otherwise brawl without the benefit of man-made tools.

\item \textbf{Recon}: scouting for danger and spotting threats.

\item \textbf{Ride}:  using animals for transportation, from riding a horse to driving a chariot to riding an elephant or dinosaur.

\item \textbf{Sail}:  controlling and using boats and ships, including navigation.

\item \textbf{Shoot}: use it as a combat skill when using ranged weaponry, whether thrown weapons, bows, pistols, combat rifles, or heavy artillery.

\item \textbf{Stab}:  fight with melee weapons or throw a hurled weapon. Maintain and identify weaponry.

\item \textbf{Sneak}: training in being unseen and unheard.

\item \textbf{Survive}: staying alive in the wilderness. This also covers \emph{outdoor} skills such as tracking, foraging, and fishing.

\item \textbf{Trade}:   Buy and sell at a profit, identify the worth of goods or treasures, deal with merchants and traders, find black-market goods and services, and know laws regarding smuggling and contraband.

\item \textbf{Work}: This skill is a catch-all for any profession that might not otherwise merit its own skill, such as a painter, lawyer, farmer, or herdsman. The precise skill it represents will vary with the PC’s background.
\end{itemize}

\section{Attribute scores}
Characteristics describe your character's general talents and abilities. They serve as basic skills for tasks which every character can perform better or worse than others.

Throw 3d6 and assign the results in order. You may then change one attribute of your choice to 14 if you wish to ensure that you character is gifted in at least one way.

\begin{description}
\item[STR (Strength)] Physical power and muscle. Governs lifting, breaking, melee combat, and feats of brute force.
\item[AGI (Agility)] Reflexes, and precision. Used for dodging, ranged attacks, acrobatics, and delicate tasks like picking locks.
\item[TGH (Toughness)] Endurance, physical resilience, and ability to withstand harm. Governs health, resistance to damage, and stamina.
\item[INT (Intelligence)] Knowledge, reasoning, and analytical skill. Governs learning, recall of information, and solving puzzles.
\item[WIL (Willpower)] Mental fortitude, determination, and force of personality. Governs magic, persuasion, resisting fear, and problem-solving.
\item[WIS (Wisdom)] Awareness, insight, and connection to the world. Used for spotting traps, sensing danger, and interpreting omens.
\end{description}

\begin{longtable}{cc}
\hline
\textbf{3d6} & \textbf{Attribute score} \\
\hline
3 & -2 \\
4-7 & -1 \\
8-13 & 0 \\
14-17 & +1 \\
18 & +2 \\
\hline
\caption{Conversion table for attribute score}
\label{table:attribute-conversion}
\end{longtable}

\section{Species} % (fold)
\label{sec:species}
In the world of Geltha, there are different species. Each one has a trait. Choose the species that suits you the most.


\section{Background}
Your past has a clear influence on you. Roll on the Family status/Family Wealth table.
Based on your habitat, choose one of the following skills:

\begin{itemize}
\item Savage Hinterlands: Notice, Survive, Exert
\item Rural Village: Animals, Ride, Work
\item Large City: Craft, Connect, Talk
\end{itemize}

\begin{longtable}{ccccc}
\hline
& \textbf{Wealth(1d12)} & \textbf{Low(1-2)} & \textbf{Middle(3-4)} & \textbf{High(5-6)} \\
\hline
Poor & 1 & Foragers & Tribal Warriors & Chieftains \\
Poor & 2 & Nomads & Wandering Shamans & Elder Mystics \\
Poor & 3 & Scavengers & Hunters & Beast Tamers \\
Poor & 4 & Outcasts & Herbalists & Spirit Guides \\
Middle & 5 & Raiders & Skilled Trackers & Clan Leaders \\
Middle & 6 & Camp Followers & Survivalists & Respected Shamans \\
Middle & 7 & Tanners & Bowyers & Tribal Council Members \\
Middle & 8 & Trappers & Spirit Channelers & Renowned Warriors \\
Rich & 9 & Mercenary Scouts & Artifact Traders & Great Chieftains \\
Rich & 10 & Adventurers & Legendary Hunters & Mystic Elders \\
Rich & 11 & Treasure Seekers & Trade Envoys & High Shamans \\
Rich & 12 & Beast Trainers & Tribal Diplomats & Warlords \\
\hline
\caption{Savage Hinterlands - Family Status/Family Wealth}
\label{table:savage-hinterlands}
\end{longtable}

% Savage Hinterlands - Hardships
\begin{longtable}{p{0.1\textwidth}p{0.9\textwidth}}
\hline
\textbf{1d20} & \textbf{Hardship} \\
\hline
1 & A childhood illness left you frail. -1 Constitution \\
2 & A hunting wound left you with a limp. -1 Dexterity \\
3 & Strange landscapes tax your stamina. -1 Constitution in foreign environments \\
4 & You struggle with foreign navigation. Roll 1d6 instead of 2d6 for Survival checks \\
5 & Endless labor distracted you from learning. -1 Intelligence \\
6 & Poor nutrition in harsh seasons stunted your growth. -1 Strength \\
7 & Exposure to harsh weather weakened your constitution. -1 Constitution \\
8 & Frequent injuries left you slower than others. -1 Dexterity \\
9 & You are easily identified as a savage native. Disguises fail quickly \\
10 & Foreign climates leave you lethargic. -1 Dexterity \\
11 & You feel uneasy with advanced tools. -1 Willpower in civilized areas \\
12 & Struggling with new customs, you make poor decisions. -1 Wisdom \\
13 & Always the runt of the group, you lack strength. -1 Strength \\
14 & A wild beast attack left you scarred. -1 Willpower \\
15 & You are clumsy with hunting tools. Roll 1d6 instead of 2d6 for Shooting checks \\
16 & A clan feud marks you for death by rivals. \\
17 & Your senses have dulled from hardship. -1 Wisdom \\
18 & A feral power sapped your vitality. Lose 3 attribute points of your choice \\
19 & Fear of spirits hinders your calm. -2 to saves against spiritual effects \\
20 & Crude tools confuse you. -2 to saves against advanced equipment effects \\
\hline
\caption{Savage Hinterlands - Hardships}
\label{table:savage-hardships}
\end{longtable}

% Rural Village - Family Status/Family Wealth
\begin{longtable}{ccccc}
\hline
& \textbf{Wealth(1d12)} & \textbf{Low(1-2)} & \textbf{Middle(3-4)} & \textbf{High(5-6)} \\
\hline
Poor & 1 & Beggars & Tenant Farmers & Wealthy Farmers \\
Poor & 2 & Vagrants & Village Laborers & Local Elders \\
Poor & 3 & Servants & Millers & Village Healers \\
Poor & 4 & Outcasts & Beekeepers & Clergy \\
Middle & 5 & Shepherds & Blacksmiths & Village Chiefs \\
Middle & 6 & Woodcutters & Bakers & Clerics or Monks \\
Middle & 7 & Fisherfolk & Carpenters & Respected Traders \\
Middle & 8 & Brickmakers & Traders & Scribes \\
Rich & 9 & Horse Breeders & Town Merchants & Landed Gentry \\
Rich & 10 & Grain Traders & Physicians & High Priests \\
Rich & 11 & Large Farm Owners & Village Leaders & Noble Representatives \\
Rich & 12 & Wine Makers & Grain Exporters & Regional Officials \\
\hline
\caption{Rural Village - Family Status/Family Wealth}
\label{table:rural-status}
\end{longtable}

% Rural Village - Hardships
\begin{longtable}{p{0.1\textwidth}p{0.9\textwidth}}
\hline
\textbf{1d20} & \textbf{Hardship} \\
\hline
1 & A village sickness weakened you. -1 Constitution \\
2 & A farming accident left you hobbled. -1 Dexterity \\
3 & Outsiders' food and drink upset your system. -1 Constitution in foreign areas \\
4 & You struggle with advanced tools. Roll 1d6 instead of 2d6 for Craft checks \\
5 & Endless farming hours left little time for study. -1 Intelligence \\
6 & A meager diet stunted your intellect. -1 Intelligence \\
7 & A famine weakened your body. -1 Strength \\
8 & Manual labor sapped your energy. -1 Constitution \\
9 & You are always recognized as a villager. Disguises fail quickly \\
10 & Foreign machinery confuses you. -1 Dexterity when operating machines \\
11 & You feel out of place in cities. -1 Willpower in urban areas \\
12 & Naivety leaves you prone to mistakes. -1 Wisdom \\
13 & Always weaker than others, you lack strength. -1 Strength \\
14 & A childhood fight left you scarred. -1 Willpower \\
15 & You're clumsy with carts and wagons. Roll 1d6 instead of 2d6 for Drive checks \\
16 & A feud with a neighboring village marks you for death. \\
17 & Your senses have dulled from village life. -1 Wisdom \\
18 & A traumatic event drained your vitality. Lose 3 attribute points of your choice \\
19 & Old tales of spirits leave you easily unnerved. -2 to saves against mental effects \\
20 & Foreign tools and gadgets confuse you. -2 to saves against magical item effects \\
\hline
\caption{Rural Village - Hardships}
\label{table:rural-hardships}
\end{longtable}

% Large City - Family Status/Family Wealth
\begin{longtable}{ccccc}
\hline
& \textbf{Wealth(1d12)} & \textbf{Low(1-2)} & \textbf{Middle(3-4)} & \textbf{High(5-6)} \\
\hline
Poor & 1 & Street Urchins & Day Laborers & Skilled Workers \\
Poor & 2 & Prostitutes & Market Vendors & Minor Clergy \\
Poor & 3 & Pickpockets & Entertainers & Guild Apprentices \\
Poor & 4 & Beggars & Artisans & Scholars \\
Middle & 5 & Petty Criminals & Shopkeepers & Guild Masters \\
Middle & 6 & Dockhands & Bakers & Respected Clergy \\
Middle & 7 & Guards & Local Traders & Town Officials \\
Middle & 8 & Street Performers & Scribes & Wealthy Merchants \\
Rich & 9 & Courtesans & Guild Leaders & City Governors \\
Rich & 10 & Moneylenders & Physicians & High Priests \\
Rich & 11 & Landed Nobility & Trade Magnates & Council Members \\
Rich & 12 & Spymasters & Foreign Diplomats & Royalty \\
\hline
\caption{Large City - Family Status/Family Wealth}
\label{table:large-city-status}
\end{longtable}

% Large City - Hardships
\begin{longtable}{p{0.1\textwidth}p{0.9\textwidth}}
\hline
\textbf{1d20} & \textbf{Hardship} \\
\hline
1 & A childhood illness left you frail. -1 Constitution \\
2 & An industrial accident left you limping. -1 Dexterity \\
3 & Pollution strains your stamina. -1 Constitution in natural settings \\
4 & Advanced systems confuse you. Roll 1d6 instead of 2d6 for Program checks \\
5 & Grueling factory work stunted your learning. -1 Intelligence \\
6 & Childhood poverty hindered your education. -1 Intelligence \\
7 & Malnutrition from slum life weakened your body. -1 Strength \\
8 & Exhausting work left you physically drained. -1 Constitution \\
9 & You are easily identified as a city native. Disguises fail quickly \\
10 & You struggle with rural environments. -1 Dexterity in non-urban areas \\
11 & You feel awkward in rural social settings. -1 Willpower in villages \\
12 & Quick decisions in new settings often fail you. -1 Wisdom \\
13 & You were always outmatched by your peers. -1 Strength \\
14 & A violent encounter left you visibly scarred. -1 Willpower \\
15 & You're terrible with vehicles. Roll 1d6 instead of 2d6 for Drive checks \\
16 & A gang rivalry makes you a target for violence. \\
17 & City noise has dulled your senses. -1 Wisdom \\
18 & A psychic event drained your vitality. Lose 3 attribute points of your choice \\
19 & Advanced tech leaves you nervous. -2 to saves against psionics \\
20 & Complex tools baffle you. -2 to saves against foreign technology effects \\
\hline
\caption{Large City - Hardships}
\label{table:large-city-hardships}
\end{longtable}


\section{Careers} % (fold)
\label{sec:careers}
Characters start off at the age of 14. Then they grow through the character generation game. Each career table presents
the various throws needed for character generation. \textbf{Note that the total sum of a character's skill can never
exceed the INT characteristic plus 6.}

\section{Career procedure}
\begin{enumerate}

  \item \textbf{Choose a Career}: Try to enteer a career. If you succeed, begin the career. If you fail, roll a 2d6 and
  pick a career accordingly to the roll. \emph{($1-2$, Vagabond, $3-11$ Commoner, $11$ Noble)}.

\item \textbf{Basic Training}: You gain Skill-0 in all six Service Skills of your career.

\item \textbf{Resolve your term of service}:
  \begin{enumerate}
    \item \textbf{Survival check}: If you fail, roll on the mishap table.
\item \textbf{Skill Table}: Roll once on one of the skill tables associated with that carrer to gain one level of a skill. Acquired skill levels are cumulative. On your first term only, roll twice and gain two skills. Note that you may roll on the Advanced Education table if your INT is over 12. 
\item \textbf{Advancement check}: Make and Advancement check. If you succeed, you gain one rank and gain an additional skill roll on any table.

\item \textbf{Rank skill}: If promoted, some ranks have skills associated with that rank. Gain that skill rank, if eligible.
\item \textbf{Aging}: If your character has completed their fourth term or later, (ages 30 and up).%, check for aging see Table~\ref{tab:aging}.

\begin{longtable}{p{0.15\textwidth} p{0.75\textwidth}} % Adjusted column widths
\hline
\textbf{2d6} & \textbf{Aging effect} \\
\hline
-4 & Reduce two physical characteristics by 3 and one mental characteristic by 2 \\ 
-3 & Reduce one physical characteristic by 3 and one mental characteristic by 2 \\
-2 & Reduce three physical characteristics by 1 \\ 
-1 & Reduce two physical characteristics by 1 \\
 0 & Reduce one physical characteristic by 1 \\ 
+1 or higher & No effect \\
\hline
\caption{Table for aging effects}
\label{tab:aging}
\end{longtable}
  \item \textbf{Events}: Roll 2d6 on the relevant Events table.

\item \textbf{Re-Enlistment}: Make a check for Re-Enlistment If you fail, you either finish your career and go to Step-4, or
    you can attempt to re-enlist in a different career(When you attempt the change add a malus equal to $-1$ times the
    previous career terms. You do not gain Skill-$0$ in the service skills.). If you succeed, you may choose to re-enlist, or you can voluntarily terminate your career. 

\item \textbf{Additional term of service}: You must retire after 7 terms of service unless you throw a natural ``12'', in which case, you must retire at the end of the 8th term of service. 
\end{enumerate}
  \item \textbf{Muster out}: Once you complete your career, roll for Mustering out benefits.  Each term of service you completed allows you one roll on one of the two Mustering out tables: Gold or Material. You gain one bonus roll if you are rank 4, two if you are rank 5, or three if you are rank 6. Character of rank 5 and 6 get a +1 bonus when rolling on the Material benefits table. No more than 3 rolls may be made on the Gold Benefit table.
\end{enumerate}
\subsection{Barbarian}
\textbf{Qualification}: STR $7+$
\textbf{Survival}: TGH $7+$
\text{Advancement}: STR $6+$
\textbf{Re-Enlistment}: $6+$ 
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.03\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Survive & Craft & Craft & 10 & Weapon \\
2 & +2 Physical & Sneak & Heal & Heal & 20 & Armor \\
3 & +2 Physical & Shoot & Lead & Lead & 30 & +1 TGH \\
4 & +2 Mental & Recon & Liaison & Liaison & 50 & Weapon \\
5 & Exert & Stab & Ride & Survive & 100 & Hound \\
6 & Any skill & Shoot & Survive & Admin & 150 & Mount \\
7 & - & - & - & - & 500 & Retainer \\
\hline
\caption{Barbarian Development, Skills, and Muster Out Benefits}
\end{longtable}


\subsection{Noble}
\textbf{Qualification}: 
\textbf{Survival}: WIL $5+$
\text{Advancement}:  $8+$
\textbf{Re-Enlistment}: $6+$ 
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Connect & Perform & Heal & 100 & Ally \\
2 & +2 Physical & Ride & Animals & Lead & 200 & Armor \\
3 & +2 Mental & Admin & Liaison & Investigation & 400 & Weapon \\
4 & +2 Mental & Sail & Admin & Sail & 600 & Steed \\
5 & Connect & Deception & Any Combat & Know & 600 & Hound \\
6 & Any skill & Connect & Any Combat & Pray & 1000 & Retainer \\
7 & - & - & - & - & 2000 & Holding \\
\hline
\caption{Noble Development, Skills, and Muster Out Benefits}
\end{longtable}
\subsection{Commoner}
\textbf{Qualification}: 
\textbf{Survival}: TGH $3+$
\text{Advancement}:  $10+$
\textbf{Re-Enlistment}: $5+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Animals & Work & Heal & 10 & +1 STR \\
2 & +2 Physical & Riding & Liaison & Lead & 10 & Hound \\
3 & +2 Physical & Craft & Connect & Liaison & 20 & Ally \\
4 & +2 Mental & Connect & Perform & Admin & 50 & Weapon \\
5 & Connect & Exert & Sail & Animals & 100 & Mount \\
6 & Any skill & Work & Animals & Pray & 100 & Freehold \\
7 & - & - & - & - & 200 & Freehold \\
\hline
\caption{Commoner Development, Skills, and Muster Out Benefits}
\end{longtable}

\subsection{Pirate}
\textbf{Qualification}: STR $7+$
\textbf{Survival}: DEX $7+$
\text{Advancement}: INT $6+$
\textbf{Re-Enlistment}: $5+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Connect & Exert & Heal & 20 & Weapon \\
2 & +2 Physical & Sail & Craft & Lead & 50 & Armor \\
3 & +2 Physical & Shoot & Deception & Perform & 100 & Weapon \\
4 & +2 Mental & Stab & Sail & Admin & 200 & +1 INT \\
5 & Any Combat & Sneak & Shoot & Know & 400 & Ally \\
6 & Any skill & Recon & Know & Investigation & 500 & +1 DEX \\
7 & - & - & - & - & 1000 & Ship \\
\hline
\caption{Pirate Development, Skills, and Muster Out Benefits}
\end{longtable}



\subsection{Priest}
\textbf{Qualification}: INT $5+$
\textbf{Survival}: WIL $5+$
\text{Advancement}: INT $8+$
\textbf{Re-Enlistment}: $6+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Pray & Pray & Lead & 10 & +1 WIL \\
2 & +2 Physical & Liaison & Survive & Deception & 10 & Holy Symbol \\
3 & +2 Mental & Lead & Craft & Know & 20 & +1 INT \\
4 & +2 Mental & Admin & Know & Heal & 50 & +1 INT \\
5 & Connect & Connect & Animals & Investigation & 50 & Ally \\
6 & Any skill & Know & Ride & Sorcery & 100 & Magic Item \\
7 & - & - & - & - & 200 & Holding \\
\hline
\caption{Priest Development, Skills, and Muster Out Benefits}
\end{longtable}

\subsection{Rogue}
\textbf{Qualification}: DEX $6+$
\textbf{Survival}: DEX $5+$
\text{Advancement}: INT $8+$
\textbf{Re-Enlistment}: $4+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Connect & Shoot & Survive & 20 & Thieves' tools \\
2 & +2 Physical & Craft & Craft & Know & 20 & Thieves' tools \\
3 & +2 Physical & Deception & Stab & Work & 100 & +1 INT \\
4 & +2 Mental & Sneak & Admin & Heal & 200 & Weapon \\
5 & Stab & Liaison & Recon & Investigation & 400 & Ally \\
6 & Any skill & Recon & Deception & Sorcery & 400 & Ally \\
7 & - & - & - & - & 1000 & Magic Item \\
\hline
\caption{Rogue Development, Skills, and Muster Out Benefits}
\end{longtable}

\subsection{Sailor}
\textbf{Qualification}: TGH $6+$
\textbf{Survival}: DEX $6+$
\text{Advancement}: INT $7+$
\textbf{Re-Enlistment}: $4+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Craft & Work & Admin & 10 & Weapon \\
2 & +2 Physical & Sail & Craft & Perform & 50 & Weapon \\
3 & +2 Physical & Shoot & Stab & Pray & 50 & +1 INT \\
4 & +2 Mental & Stab & Connect & Heal & 100 & Magic Item \\
5 & Exert & Exert & Sail & Survive & 100 & Ally \\
6 & Any skill & Recon & Work & Work & 200 & Ally \\
7 & - & - & - & - & 200 & Ship \\
\hline
\caption{Sailor Development, Skills, and Muster Out Benefits}
\end{longtable}

\subsection{Scholar}
\textbf{Qualification}: INT $6+$
\textbf{Survival}: TGH $4+$
\text{Advancement}: STR $9+$
\textbf{Re-Enlistment}: $5+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Admin & Alchemy & Pray & 10 & +1 INT \\
2 & +2 Physical & Know & Admin & Perform & 50 & +2 INT \\
3 & +2 Mental & Heal & Heal & Magic & 50 & Ally \\
4 & +2 Mental & Liaison & Know & Heal & 100 & Instruments \\
5 & Connect & Investigation & Craft & Know & 100 & Holding \\
6 & Any skill & Know & Carousing & Work & 200 & Instruments \\
7 & - & - & - & - & 250 & Magic Item \\
\hline
\caption{Scholar Development, Skills, and Muster Out Benefits}
\end{longtable}

\subsection{Shaman}
\textbf{Qualification}: INT $9+$
\textbf{Survival}: WIL $7+$
\text{Advancement}: INT $6+$
\textbf{Re-Enlistment}: $6+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Pray & Shoot & Heal & 10 & Weapon \\
2 & +2 Physical & Magic & Artifice & Perform & 10 & Spell \\
3 & +2 Mental & Survive & Recon & Know & 20 & Focus \\
4 & +2 Mental & Know & Ride & Know & 20 & Magic Item \\
5 & Connect & Heal & Liaison & Pray & 50 & Spell \\
6 & Any skill & Animals & Lead & Magic & 50 & Mount \\
7 & - & - & - & - & 100 & Talisman \\
\hline
\caption{Shaman Development, Skills, and Muster Out Benefits}
\end{longtable}

\subsection{Soldier}
\textbf{Qualification}: TGH $8+$
\textbf{Survival}: STR $6+$
\text{Advancement}: INT $7+$
\textbf{Re-Enlistment}: $6+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Craft & Craft & Heal & 10 & Weapon \\
2 & +2 Physical & Shoot & Lead & Know & 50 & Hound \\
3 & +2 Physical & Shoot & Sneak & Work & 50 & Weapon \\
4 & +2 Mental & Stab & Liaison & Lead & 100 & Armor \\
5 & Exert & Recon & Survive & Perform & 100 & Mount \\
6 & Any skill & Ride & Shoot & Know & 200 & Armor \\
7 & - & - & - & - & 200 & Steed \\
\hline
\caption{Soldier Development, Skills, and Muster Out Benefits}
\end{longtable}

\section{Sorcerer} 
\textbf{Qualification}: INT $9+$
\textbf{Survival}: WIL $8+$
\text{Advancement}: INT $5+$
\textbf{Re-Enlistment}: $6+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Know & Pray & Know & 20 & +1 INT \\
2 & +2 Physical & Magic & Deception & Pray & 50 & Magic \\
3 & +2 Mental & Deception & Perform & Lead & 50 & Magic \\
4 & +2 Mental & Investigation & Heal & Work & 100 & Familiar \\
5 & Deception & Know & Liaison & Heal & 100 & Magic \\
6 & Any skill & Magic & Craft & Know & 200 & Magic Item \\
7 & - & - & - & - & 300 & Talisman \\
\hline
\caption{Sorcerer Development, Skills, and Muster Out Benefits}
\end{longtable}
\section{Wanderer}
\textbf{Qualification}: 
\textbf{Survival}: TGH $7+$
\text{Advancement}: INT $6+$
\textbf{Re-Enlistment}: $6+$
\subsubsection{Skills and Muster Out Benefits}
\begin{longtable}{p{0.05\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth} p{0.15\textwidth}}
\hline
\textbf{1d6} & \textbf{Personal Dev.} & \textbf{Service Skills} & \textbf{Specialist Skills} & \textbf{Advanced Education} & \textbf{Money (Gold)} & \textbf{Material Benefits} \\
\hline
1 & +1 Any Stat & Survive & Sail & Perform & 10 & Weapon \\
2 & +2 Physical & Sneak & Craft & Work & 20 & Hound \\
3 & +2 Physical & Stab & Perform & Lead & 30 & Ally \\
4 & +2 Mental & Ride & Animals & Know & 30 & Weapon \\
5 & Perform & Sneak & Athletics & Heal & 100 & Ally \\
6 & Any skill & Recon & Deception & Know & 100 & Mount \\
7 & - & - & - & - & 200 & Mount \\
\hline
\caption{Wanderer Development, Skills, and Muster Out Benefits}
\end{longtable}

\end{document}